---
title: "Bonanza Creek Snowshoe Hairs Mini Report"
author: "Jaxon Stuhr"
date: "11/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(kableExtra)
library(lubridate)
```

```{r}
# read in data
bonanza_hares_raw = read_csv(here("data", "bonanza_hares.csv"))
```


## Introduction

## Data and Analyses

## Juvenile Hare Initial Investigation

### Annual juvenile hare trap counts
```{r}
# build working dataset
juv_bonanza_hares = bonanza_hares_raw %>% 
  # filter for juveniles
  filter(age == "j") %>% 
  # lubridate date
  mutate(date = mdy(date)) %>% 
  # add yr column
  mutate(year = year(date))
```
```{r}
# ggplot w x = year
ggplot(juv_bonanza_hares, aes(x = year)) + 
  # hist w 14 bins
  geom_histogram(binwidth = 1, fill = "Dark Green") + 
  # build x axis labels
  scale_x_continuous(
    name="Year", 
    limits=c(1998, 2013), 
    breaks=c(1999,2001, 2003, 2005, 2007, 2009, 2011)) +
  # add theme
  theme_linedraw() +
  # add labels
  labs(x = "Year", 
       y = "Juvenile Hair Trap Counts", 
       title = "Juvenile Hair Trap Counts by Year") +
  # center title
  theme(plot.title = element_text(hjust = 0.5))
```

Fig 1: Juvenile Hair Trap Counts at Bonanza Creek Experimental Yeared for Yesars 1999 - 2012
```{r}
# summary stats for analyses
juv_trap_counts_summary = juv_bonanza_hares %>% 
  # group by yr
  group_by(year) %>% 
  # count
  summarize(
    count = n()
  )
# mean, median, max, min juvenile traps for 1999 - 2012
mean_traps = mean(juv_trap_counts_summary$count)
median_traps = median(juv_trap_counts_summary$count)
max_traps = max(juv_trap_counts_summary$count)
min_traps = min(juv_trap_counts_summary$count)
```
#### Analyses:

(((add this)))

### Visualize juvenile hare weights

### Juvenile weight comparison (male & female snowshoe hares)

### Relationship between juvenile weight & hind foot length

## Summary

## References

