---
title: "Bonanza Creek Snowshoe Hairs Mini Report"
author: "Jaxon Stuhr"
date: "11/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(kableExtra)
library(lubridate)
library(effsize)
library(ggpubr)
library(broom)
```

```{r}
# read in data
bonanza_hares_raw = read_csv(here("data", "bonanza_hares.csv"))
```


## Introduction

## Data and Analyses

## Juvenile Hare Initial Investigation

### Annual juvenile hare trap counts
```{r}
# build working dataset
juv_bonanza_hares = bonanza_hares_raw %>% 
  # filter for juveniles
  filter(age == "j") %>% 
  # lubridate date
  mutate(date = mdy(date)) %>% 
  # add yr column
  mutate(year = year(date)) %>% 
  # add full site name column
  mutate(full_site = case_when(
    grid == "bonrip" ~ "Bonanza Riparian",
    grid == "bonbs" ~ "Bonanza Black Spruce",
    grid == "bonmat" ~ "Bonanza Mature"
  ))
```
```{r}
# ggplot w x = year
ggplot(juv_bonanza_hares, aes(x = year)) + 
  # hist w 14 bins
  geom_histogram(binwidth = 1, fill = "Dark Green") + 
  # build x axis labels
  scale_x_continuous(
    name="Year", 
    limits=c(1998, 2013), 
    breaks=c(1999,2001, 2003, 2005, 2007, 2009, 2011)) +
  # add theme
  theme_linedraw() +
  # add labels
  labs(x = "Year", 
       y = "Juvenile Hare Trap Counts")   
      # title = "Juvenile Hare Trap Counts by Year") +
  # center title
  #theme(plot.title = element_text(hjust = 0.5))
```

Fig 1: Juvenile hare trap counts at Bonanza Creek Experimental Forest for years 1999 - 2012
```{r}
# summary stats for analyses
juv_trap_counts_summary = juv_bonanza_hares %>% 
  # group by yr
  group_by(year) %>% 
  # count
  summarize(
    count = n()
  )
# mean, median, max, min juvenile traps for 1999 - 2012
mean_traps = mean(juv_trap_counts_summary$count)
median_traps = median(juv_trap_counts_summary$count)
max_traps = max(juv_trap_counts_summary$count)
min_traps = min(juv_trap_counts_summary$count)
```
#### Analyses:

The maximum number of hare trappings in a given year was `r max_traps` and occurred in 1999, while the minimum was zero and occurred in both 2002 and 2009. The mean annual trappings was `r mean_traps` and the median was `r median_traps`. The disparity between these two is due to the huge difference in trappings between 1999 and any other year, thus the 1999 trappings pulled up the mean. There's a slight trend of decreasing hare trappings, however, this depends heavily on the 1999 data point which is a stark outlier. To better understand this trend one would have to account for the annual effort, or number of traps laid and how often. With the given data it is impossible to tell if differences between years are due to changing hare populations or changing effort put in to trapping.

### Visualize juvenile hare weights

```{r}
# ggplot w x = sex, y = weight
ggplot(juv_bonanza_hares, aes(x = sex, y = weight, color = sex)) +
  # jitter plot
  geom_jitter() + 
  # group by site name
  facet_wrap(~ full_site) +
  # add labels
  labs(x = "Sex", 
       y = "Weight [grams]")+ 
       #title = "Juvenile Hare Weights at Bonanza Trapping Locations") +
  # rename x-axis labels
  scale_x_discrete(labels = c("Female","Male","NA")) + 
  # remove legend
  guides(color = FALSE) 
  # center title
  #theme(plot.title = element_text(hjust = 0.5))
```

Fig 2: Juvenile hare weights [g] at three Bonanza sites sorted by site and sex for hares trapped between 1999 and 2012
```{r}
# summary stats for analyses
juv_weights_sex_site = juv_bonanza_hares %>% 
  # group by sex and site
  group_by(sex, full_site) %>% 
  # add mean, sd, sample
  summarise(
    mean_weight = round(mean(weight, na.rm = TRUE),1),
    sd_weight = round(sd(weight, na.rm = TRUE),1),
    sample_size = n()
  ) 
```
#### Analyses:

At all sites male hares weighed on average `r juv_weights_sex_site$mean_weight[4] - juv_weights_sex_site$mean_weight[1]` to `r juv_weights_sex_site$mean_weight[5] - juv_weights_sex_site$mean_weight[2]` grams more than females. Hares at the Bonanaza black spruce site also weighed more on average that at the mature and riparian sites, but there was not a large variation from site to site or by sex.

### Juvenile weight comparison (male & female snowshoe hares)

```{r}
# summary stats for analyses
juv_weights_sex = juv_bonanza_hares %>% 
  # group by sex and site
  group_by(sex) %>% 
  # add mean, sd, sample
  summarise(
    mean_weight = round(mean(weight, na.rm = TRUE),1),
    sd_weight = round(sd(weight, na.rm = TRUE),1),
    sample_size = n()
  ) 

juv_weights_sex %>%
  # change labels for male, female
  mutate(sex = case_when(
    sex == "f" ~ "Female",
    sex == "m" ~ "Male"
  )) %>% 
  # build table w kbl extra, rename columns
  kbl(caption = "Table 1: Juvenile Hare Weights for Bonanza Sites", col.names = c("Sex", "Mean Weight [g]", "Standard Deviation [g]", "Sample Size")) %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

```{r, include = FALSE}
# plot exploratory histograms and male and female weights to check for high-level normality 
male_weights = juv_bonanza_hares %>% 
  filter(sex == "m")
ggplot(male_weights, aes(x = weight)) + 
  geom_histogram(bins = 10)
female_weights = juv_bonanza_hares %>% 
  filter(sex == "f")
ggplot(female_weights, aes(x = weight)) + 
  geom_histogram(bins = 10)
```

```{r}
# percent difference of weights
p_diff = 200 *
  (juv_weights_sex$mean_weight[2] - juv_weights_sex$mean_weight[1]) /
  (juv_weights_sex$mean_weight[2] + juv_weights_sex$mean_weight[1])
```


```{r}
# build vectors for cohen's d, t-test using "pull"
m_weights_vec = male_weights %>% # male weights as vector
  pull(weight)
f_weights_vec = female_weights %>% # female weights as vector
  pull(weight)
```

```{r}
# cohen's d effect size 
weights_cohen = cohen.d(m_weights_vec, f_weights_vec, na.rm = TRUE)
```

```{r}
# t-test, unpaired, 2 sample
weights_t = t.test(m_weights_vec, f_weights_vec)
```

#### Analyses

On average, trapped male juvenile hares weighed `r juv_weights_sex$mean_weight[2] - juv_weights_sex$mean_weight[1]` grams, or `r round(p_diff,1)`%, more than trapped female juvenile hares. Examining the effect of hare sex on weight with a t-test, we find that there is less than a 1% chance of randomly drawing samples from populations with the same mean that are at least as different as the male and female hairs were. The effect, however, was small (Cohen's *d* = `r round(weights_cohen$estimate, 2)`).

### Relationship between juvenile weight & hind foot length

```{r}
# linear model of hare weight as a function of hind foot length
hare_lm = lm(weight ~ hindft, juv_bonanza_hares)
# glance to acess outputs
hare_lm_out = glance(hare_lm)
# summary to view
# summary(hare_lm)
```

```{r}
# pearson's r to find correlation between bodymass, hind foot length
hares_cor = cor.test(juv_bonanza_hares$weight, juv_bonanza_hares$hindft)
```

```{r}
# ggplot of hind foot length vs weight
ggplot(juv_bonanza_hares, aes(x = hindft, y = weight)) +
  # scatter plot
  geom_point(color = "blue") + 
  # add linear regression fitted line
  geom_smooth(method = "lm",
              color = "red",
              size = 0.5,
              fill = "gray10",
              alpha = 0.5) +
  theme_light() + 
  # add labels
  labs(x = "Hind foot length [mm]", y = "Weight [g]") +
  # add equation of line at given position
  ggpubr::stat_regline_equation(label.x = 70, label.y = 1250)
```

Fig 3: Juvenile hare weights [g] as a function of hind foot length [mm] at three Bonanza sites, hares trapped between 1999 and 2012 with fitted linear regression

#### Analysis

The fitted linear regression model states that for every 1mm increase in hind foot length, the hares' weight increase by `r round(hare_lm$coefficients[2],1)` grams. The fitted line had an R^2^ = `r round(hare_lm_out[2],2)` meaning only `r 100*round(hare_lm_out[2],2)`% of the variance in weight could be explained by changes in hind foot length. From a Pearson's *r* value of `r round(hares_cor$estimate,2)` we see that the variables (hind foot length and weight) are moderately correlated. The assumption of homoscedasticity, which is required for linear regression, appears to not apply as we can see in the scatter plot that variance in weight increases as hind foot length gets bigger. 

## Summary

## References

